Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-20 manuscript/appC.Rmd
adds more on the pasm recruitment booms

Appendix C: Effect of recruitment on IPM uncertainty
=====================================================
In Fig. 4 of the main text where we show the uncertainty around our models' forecasts to a 1% change in climate covariates, the obvious outlier is *Pascopyrum smithii* (PASM). The uncertainty around PASM forecasts from the integral projection model is orders of magnitude larger than for any other species, and is even larger than the uncertainty around the quad-based model forecasts. It turns out that the recruitment parameters we estimate lead to overcompensating density-dependence that is amplified when those parameters are allowed to vary. When we allow all parameters to vary except those associated with recruitment, uncertainty around PASM decreases to levels similar to those associated with other species (Fig. C1). So detailed demographic data is not a cure-all, and can sometimes lead to interesting dynamics when parameter uncertainty is considered.

>>>>

Appendix C: Effect of recruitment on IPM uncertainty
=====================================================
Seems to be due to the climate effects. Are they really more uncertain than the other species -- or is it more that it is such low cover so boom years are more influential? I think it is because average recruits are very high relative to other species, so then tack on occassional big years due to climate draws, and you get booms with subsequent low survival busts.

In Fig. 4 of the main text where we show the uncertainty around our models' forecasts to a 1% change in climate covariates, the obvious outlier is *Pascopyrum smithii* (PASM). The uncertainty around PASM forecasts from the integral projection model is orders of magnitude larger than for any other species, and is even larger than the uncertainty around the quad-based model forecasts. It turns out that the estimated dispersion parameter for the negative binomial distribution is much higher for PASM than for other species (Appendix B). This is because PASM has higher recruitment rates than the other species and the relationship between number of recruits and quadrat cover (see main text) is weak. So, when we draw parameters from the MCMC chain for our simulations, even larger values for the dispersion parameter are possible. Thus, there can be very large "boom" years where recruitment is high. When we allow all parameters to vary except those associated with recruitment, uncertainty around PASM decreases to levels similar to those associated with other species (Fig. C1). We also performed simulations where all recruitment parameters except the dispersion parameter were allowed to vary. Results from those simulations show the effect of the dispersion parameter on boom and bust years (Fig. C2).

### 2015-07-20 manuscript/manuscript.Rmd
adds more on the pasm recruitment booms

### Model forecasts
Forecasts based on 1% climate changes were extremely uncertain when we considered model error and parameter uncertainty (Fig. 6). As expected based on model validation (Table 1), QBM projections were more uncertain than IPM projections for all species except *P. smithiii* (Fig. 6). IPM forecasts for *P. smithiii* were very uncertain due to overcompensating density dependence and nonlinearities recruitment that get amplified by parameter uncertainty (Appendx C). Neither model was capable of making forecasts of proportional cover change distinguishable from 0 (i.e., no change in population state) when we included model error and parameter uncertainty.

>>>>

### Model forecasts
Forecasts based on 1% climate changes were extremely uncertain when we considered model error and parameter uncertainty (Fig. 6). As expected based on model validation (Table 1), QBM projections were more uncertain than IPM projections for all species except *P. smithiii* (Fig. 6). IPM forecasts for *P. smithiii* were very uncertain due to a large dispersion parameter ($\zeta$ in Eq. 6) that can lead to boom years with very large recruitment and subsequent busts due to low survival of so many young plants (Appendx C). Neither model was capable of making forecasts of proportional cover change distinguishable from 0 (i.e., no change in population state) when we included model error and parameter uncertainty.

### 2015-07-20 manuscript/manuscript.tex
adds more on the pasm recruitment booms

Forecasts based on 1\% climate changes were extremely uncertain when we
considered model error and parameter uncertainty (Fig. 6). As expected
based on model validation (Table 1), QBM projections were more uncertain
than IPM projections for all species except \emph{P. smithiii} (Fig. 6).
IPM forecasts for \emph{P. smithiii} were very uncertain due to
overcompensating density dependence and nonlinearities recruitment that
get amplified by parameter uncertainty (Appendx C). Neither model was
capable of making forecasts of proportional cover change distinguishable
from 0 (i.e., no change in population state) when we included model
error and parameter uncertainty.

>>>>

Forecasts based on 1\% climate changes were extremely uncertain when we
considered model error and parameter uncertainty (Fig. 6). As expected
based on model validation (Table 1), QBM projections were more uncertain
than IPM projections for all species except \emph{P. smithiii} (Fig. 6).
IPM forecasts for \emph{P. smithiii} were very uncertain due to a large
dispersion parameter ($\zeta$ in Eq. 6) that can lead to boom years with
very large recruitment and subsequent busts due to low survival of so
many young plants (Appendx C). Neither model was capable of making
forecasts of proportional cover change distinguishable from 0 (i.e., no
change in population state) when we included model error and parameter
uncertainty.
